---
title: "Diamonds Exploration (9/28)"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading Data

```{r}
mydiamonds=read.csv("diamonds.csv")
```

# Dimension and names of variables

```{r}
dim(mydiamonds)
nrow(mydiamonds)
ncol(mydiamonds)
```

There are `r nrow(mydiamonds)` rows in our dataset. 

```{r}
colnames(mydiamonds)
mydiamonds=mydiamonds[,2:11]
```

After knocking off the first column, I am now left with `r ncol(mydiamonds)` columns.

# saving my truncated file

```{r}
save(mydiamonds,file="mydiamondsset2.rda")
```

# data structure

```{r}
str(mydiamonds)
```

```{r}
summary(mydiamonds$depth)
```

Number of levels of a factor variable

```{r}
nlevels(mydiamonds$clarity)
nlevels(mydiamonds[,c("clarity","cut")])
```

```{r}
fairdiamonds=mydiamonds[mydiamonds$cut=="Fair",]
```

```{r}
levels(fairdiamonds$cut)
summary(fairdiamonds$cut)
```

# refactoring after a subset of factor variable
  -after making "fairdiamonds" only fair cut, r just thinks the other four levels are blank and have 0 observations. I need to remove the other levels
  
  
```{r}
fairdiamonds$cut=factor(fairdiamonds$cut)
summary(fairdiamonds$cut)
```
